---
title: "Tidy Text Mining of David Bowie's Lyrics"
author: "Nic Crane"
output:
  html_document: 
    keep_md: true
---

After hearing about **tidytext** from colleagues who went to the RStudio conference last month, I decided to try it out.  If you want to try out tidy text 
mining for yourself, check out the official website at <http://tidytextmining.com>.

Legendary performer David Bowie would have been 70 earlier this month, and as a 
tribute to his impressive career, I decided to take a look at his lyrics.

First, I used the **rvest** package to scrape all the lyrical content from all
albums on DavidBowie.com.  Only albums from before 2016 which had lyrics present
on this website and weren't re-releases were included (i.e. the eponymous album 
of 1969 was omitted as it was re-released under the title "Space Oddity").

```{r, echo = FALSE}
suppressMessages({
suppressWarnings({
  
library(dplyr)
library(tidytext)
library(data.table)
library(knitr)
  
})

load("bowieData.Rda")
load("bowieLyrics.Rda")
bowieData <- select(bowieData, Album, Year, BPI)
})

```
Let's take a look at the structure of the lyrics dataset.
```{r}
str(bowieLyrics)
```

The dataset contains a character string for each track, with the lyrics
for the whole song.  It looks good, but it's a bit of a mess, and definitely
needs more work before we can use it for any kind of analysis.

The first job is to tidy up the dataset.  The `unnest_tokens` function in the 
tidytext package makes easy work of this for us.

```{r cache = TRUE}
# tidy data so it has one-token-per-document-per-row format
unnested <- unnest_tokens(bowieLyrics, word, lyrics)
kable(head(unnested, 5))
```

Looking good, although these lyrics are displayed in the order in which they 
occur.  Let's sort them and take a look at what the most common words are.

```{r}
# What are the most common words? 
count(unnested, word, sort=TRUE)
```

That's great, but not so interesting.  Let's remove words like 'the' and 'and', 
and take another look. 

```{r}
data("stop_words")
unnested <- unnested %>% anti_join(stop_words)

# Cool, let's try again!
count(unnested, word, sort=TRUE)
```

OK, this is reminding me a lot more of the Bowie songs that I know and love!

I bet this would look better as a word cloud though...

```{r, message=FALSE}
library(wordcloud)

# Some interesting results - one of which is the word "ch" as in "ch ch changes"!
unnested %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 120, colors="purple4"))

```

Finally, let's take a quick look at words used to express sentiments in Bowie's lyrics.

The `get_sentiments` function allows us to retrieve sentiment lexicons.  They're in
a tidy format so can easily be joined with our tidy text data.

First, we join the datasets and create a word cloud of only the positive words.

```{r, message=FALSE}
# Sentiment analysis of lyrical content
bing <- get_sentiments("bing")

byEmotion <- unnested %>%
  inner_join(bing)

byEmotion %>%
  filter(sentiment == "positive") %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 50, colors="red"))
```

By far the most common positive word in David Bowie's lyrics was 'love'.

Now let's take a look at the negative words.

```{r}
byEmotion %>%
  filter(sentiment == "negative") %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 50, colors="black"))
```

I think I prefer the first word cloud!

This was just a quick look at the tidytext package, but we're a big fan of how
easy it is to quickly analyse text data.  Check out <http://tidytextmining.com/> 
to see what else this package can do!


This blogpost (with code) can also be found on [GitHub](https://github.com/MangoTheCat/blog_bowie_tidytext).
